Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-03-01 additional_files/main_text.tex
Invert the Methods and Results section; provide more context at the beginning of the results to compensate

Alternative approaches that allow one to assess a large spectrum
of variants in GWAS without genotyping SVs at population scale could be useful.
\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
represent such an alternative method. This approach relies on the presence or
absence of \emph{k}-mers observed in sequence reads instead of genotypes for use in
GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
population under study, this table can be used in association analyses
similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
that they can act as a molecular signature for any type of variant as long as
these variants result in presence/absence of \emph{k}-mers of a given length.
Another advantage is that \emph{k}-mers are not tied to a specific genomic
location and can thus be used to query associations between genotype and
phenotype at genomic locations that are not found in the reference genome.  A
few studies have applied this \emph{k}-mer-based GWAS approach
\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
such approaches have not been widely adopted yet.

@@

We analyzed a total of 13 traits using three different GWAS approaches, one
based on SNP and indel genotypes (Platypus), one based on SV genotypes (Paragraph),
and one based on the presence/absence of \textit{k}-mers.
Because two or three different GWAS analyses were conducted for some of the traits, a
total of 22 GWAS analyses were performed using each of the all approaches.
Given the large volume of data that this represents, we will focus here on the
most noteworthy results. Readers interested in detailed results for loci not
discussed here are referred to Supplemental data. In particular, table \ref{signals_table}
provides a summary of the \textit{p}-values obtained by each approach at each
of the loci considered in this study.

>>>>

Alternative approaches that allow one to assess a large spectrum
of variants in GWAS without genotyping SVs at population scale could be useful.
\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
represent such an alternative method. This approach relies on the presence or
absence of \emph{k}-mers observed in sequence reads instead of genotypes for use in
GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
population under study, this table can be used in association analyses
similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
that they can act as a molecular signature for any type of variant as long as
these variants result in presence/absence of \emph{k}-mers of a given length.
Another advantage is that \emph{k}-mers are not tied to a specific genomic
location and can thus be used to query associations between genotype and
phenotype at genomic locations that are not found in a reference genome.  A
few studies have applied this \emph{k}-mer-based GWAS approach
\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
such approaches have not been widely adopted yet.

@@

We analyzed a total of 13 traits using three GWAS approaches that differed in
the genotypes used. The first approach used SNP and indel
genotypes called by Platypus. Despite both SNPs and indels being included in
this approach, we will refer to this approach as SNP-based for simplicity. Our
SV-based workflow used various sources of data for SV discovery and a single
program (Paragraph) for genotyping SVs at population scale. Finally, the third
GWAS approach relied on the presence/absence of \textit{k}-mers as genotypes.
Because of the large number of \textit{k}-mers analyzed, we only report results
(e.g. in Manhattan plots) for those that were considered significantly
associated with traits of interest.

Since two or three different GWAS analyses were conducted for some of the traits, a
total of 22 GWAS analyses were performed using each of the three approaches.
Given the large volume of data that this represents, we focus here on the
most noteworthy results. Readers interested in results at loci that are not
discussed here are referred to Supplemental data. In particular, table \ref{signals_table}
provides a summary of the \textit{p}-values obtained by each approach at each
of the loci considered in this study.
